Statically Typed String Sanitation Inside a Python

### 2014-08-13 sanitation-psp14/sanitation-psp14-short.tex
typos

Our type system for regular string types, $\lambdas$, is defined as an exntenstion to a simply-typed lambda calculus.
Opeartions on strings are given types according to corresponding, type-level
computations on regular expressions. A concatenation operation is defined in 
order to demosntrate this principle, but the premier operation of our system
is \emph{replacement}, which captures a rich set of input sanitation techniques
including escaped characters and encodings. In addition to defining this system,
we present a correctness-preserving compilation of $\lambdas$ into an underlying
language with ordinary strings as a regular expression library.

@@

Summarily, we present a simple type system extension which ensures the absence
of input sanitation vulnerabilities by statically checking the correctness and 
correct use of input sanitation algorithms. This approach is \emph{composable}
in the sense that it is a conservative extension. This approach is \emph{complementary}
to exsiting input sanitation techniques that use regular expressions for input santiation.

@@

The input sanitation problem is well-understood. There exist a large number of techniques and technologies, proposed by both practitioners and researchers, for preventing injection-style attacks. In this section, we explain how our approach to the input sanitation problem differs from each of these approaches. More important than these differences, however, is our more general assertion that language extensibility is a promising approach toward consideration of security goals in programming lanugage design.

@@

The Wyvern programming language provides a general framework for composing lanugage extensions \cite{tsl}\cite{wyvern}.
Our work identifies one particular extension, and is therefore complementary to Wyvern and related
work on extensible programming languages.
We are also unaware of any extensible programming languages which emphasize applications to security concerns.

Incorporating regular expressions into the type system is not novel. The XDuce system \cite{HosoyaVouillonPierce2000ICFP,HosoyaPierce2002} checks XML documents against schemas using regular expressions. 
Similarly, XHaskell \cite{xhaskell} focuses on XML documents.
We differ from this and related work in at least three ways:
\begin{itemize}
  \item Our system is defined within an extensible type system.
  \item We demonstrate that regular expression types are applicable to the web security domain, whereas previous work on regular expression types focused on XML schemas.
  \item Although our static replacement operation is definable in some languages with regular expression types, we are the first to expose this operation and connect the semantics of regular language replacement with the semantics of string substitution via a type safety and compilation correctness argument.
\end{itemize}

@@

The premier operation for manipulating strings in $\lambdas$ is string substitution, 
which is a familiar operation to any programmer who has used regular expressions.
The replacement operation replaces all instances of a pattern in one string with
another string; for instance, $\lsubst{a|b}{a}{c} = c$.
In order to compute the type resulting from aubstitution, we also need to compute
the result of replacing one language with another inside a given language.
Finally, just for convienance, we provide a coerce operation. The introduction of
coercion requires handling of runtime errors.

@@

A brief outline of this section follows:
\begin{itemize}
  \item Page 3 contains a definition of $\lambdas, \lambdap$ and the translation from $\lambdas$ to $\lambdap$. Grammar follows immediately at the top of page 4.
  \item In \S 2.1 we state some properties about regular expressions which are needed in our correctness proofs.
   \item In \S 2.2 we prove type safety for $\lambdap$ as well as both type safety and correctness for $\lambdas$.
  \item In \S 2.3 we prove that translation preserves the correctness reesult about $\lambdas$.
\end{itemize}

@@

The system $\lambdap$ is a stright-forward extension of a simply typed lambda claculus
with a string type and a regular epxression tpye. We include two operations which are available in the regular expression
library of any modern programming language. The check operation ensures that an expression
recognizes a string, and the replace operation is string replacement.

@@

\begin{defn}[$\tt{lsubst}$]
The replation $\lsubst{r}{s_1}{s_2} = s$ produces a string $s$ in which all substrings of $s_1$ matching $r$ are replaced with $s_2$.
\end{defn}

@@

\begin{lem}{Substitution Correspondence.} \label{thm:substcorrespondence}
  If $s_1 \in \lang{r_1}$ and $s_2 \in \lang{r_2}$ then $\lsubst{r}{s_1}{s_2} \in \lang{\lreplace{r}{s_1}{s_2}}$.
\end{lem}
\begin{proof}
The theorem follows from the definitions of lsubst and lreplace; note that language substitutions over-approximate string substititons.
\end{proof}

@@

The result of applying sanitize to input is apped to two functions which
construct a safe query and safe output.
The arguments to the result and output construction functions constitute \emph{specifications}.
In the case of \textsf{results_query}, this specification ensures that user input is always interpreted
as a string literal by the SQL server.
In the case of \textsf{results_div}, this specification ensures that user input does 
not contain any HTML tags, which is a conservative but effective policy for preventing
XSS attacks.

@@

The \textsf{string_in} type has an indexing regular expression \texttt{idx}.
Our translation is defined by the \textsf{trans_} methods while the \textsf{syn_}
methods define our type checker. The atlang type checker or translater defer
to these methods whenever an expression of type \textsf{string_in} is encountered.

>>>>

Our type system for regular string types, $\lambdas$, is defined as an exntension to a simply-typed lambda calculus.
Operations on strings are given types according to corresponding, type-level
computations on regular expressions. A concatenation operation is defined in 
order to demonstrate this principle, but the premier operation of our system
is \emph{replacement}, which captures a rich set of input sanitation techniques
including escaped characters and encodings. In addition to defining this system,
we present a correctness-preserving compilation of $\lambdas$ into an underlying
language with ordinary strings as a regular expression library.

@@

Summarily, we present a simple type system extension which ensures the absence
of input sanitation vulnerabilities by statically checking the correctness and 
correct use of input sanitation algorithms. This approach is \emph{composable}
in the sense that it is a conservative extension. This approach is \emph{complementary}
to existing input sanitation techniques that use regular expressions for input sanitation.

@@

The input sanitation problem is well-understood. There exist a large number of techniques and technologies, proposed by both practitioners and researchers, for preventing injection-style attacks. In this section, we explain how our approach to the input sanitation problem differs from each of these approaches. More important than these differences, however, is our more general assertion that language extensibility is a promising approach toward consideration of security goals in programming language design.

@@

The Wyvern programming language provides a general framework for composing language extensions \cite{tsl}\cite{wyvern}.
Our work identifies one particular extension, and is therefore complementary to Wyvern and related
work on extensible programming languages.
We are also unaware of any extensible programming languages which emphasize applications to security concerns.

Incorporating regular expressions into the type system is not novel. The XDuce system \cite{HosoyaVouillonPierce2000ICFP,HosoyaPierce2002} checks XML documents against schema using regular expressions. 
Similarly, XHaskell \cite{xhaskell} focuses on XML documents.
We differ from this and related work in at least three ways:
\begin{itemize}
  \item Our system is defined within an extensible type system.
  \item We demonstrate that regular expression types are applicable to the web security domain, whereas previous work on regular expression types focused on XML schema.
  \item Although our static replacement operation is definable in some languages with regular expression types, we are the first to expose this operation and connect the semantics of regular language replacement with the semantics of string substitution via a type safety and compilation correctness argument.
\end{itemize}

@@

The premier operation for manipulating strings in $\lambdas$ is string substitution, 
which is a familiar operation to any programmer who has used regular expressions.
The replacement operation replaces all instances of a pattern in one string with
another string; for instance, $\lsubst{a|b}{a}{c} = c$.
In order to compute the type resulting from substitution, we also need to compute
the result of replacing one language with another inside a given language.
Finally, just for convenience, we provide a coerce operation. The introduction of
coercion requires handling of runtime errors.

@@

A brief outline of this section follows:
\begin{itemize}
  \item Page 3 contains a definition of $\lambdas, \lambdap$ and the translation from $\lambdas$ to $\lambdap$. Grammar follows immediately at the top of page 4.
  \item In \S 2.1 we state some properties about regular expressions which are needed in our correctness proofs.
   \item In \S 2.2 we prove type safety for $\lambdap$ as well as both type safety and correctness for $\lambdas$.
  \item In \S 2.3 we prove that translation preserves the correctness result about $\lambdas$.
\end{itemize}

@@

The system $\lambdap$ is a straight-forward extension of a simply typed lambda calculus
with a string type and a regular expression tpye. We include two operations which are available in the regular expression
library of any modern programming language. The check operation ensures that an expression
recognizes a string, and the replace operation is string replacement.

@@

\begin{defn}[$\tt{lsubst}$]
The relation $\lsubst{r}{s_1}{s_2} = s$ produces a string $s$ in which all substrings of $s_1$ matching $r$ are replaced with $s_2$.
\end{defn}

@@

\begin{lem}{Substitution Correspondence.} \label{thm:substcorrespondence}
  If $s_1 \in \lang{r_1}$ and $s_2 \in \lang{r_2}$ then $\lsubst{r}{s_1}{s_2} \in \lang{\lreplace{r}{s_1}{s_2}}$.
\end{lem}
\begin{proof}
The theorem follows from the definitions of lsubst and lreplace; note that language substitutions over-approximate string substitutions.
\end{proof}

@@

The result of applying sanitize to input is appended to two functions which
construct a safe query and safe output.
The arguments to the result and output construction functions constitute \emph{specifications}.
In the case of \textsf{results_query}, this specification ensures that user input is always interpreted
as a string literal by the SQL server.
In the case of \textsf{results_div}, this specification ensures that user input does 
not contain any HTML tags, which is a conservative but effective policy for preventing
XSS attacks.

@@

The \textsf{string_in} type has an indexing regular expression \texttt{idx}.
Our translation is defined by the \textsf{trans_} methods while the \textsf{syn_}
methods define our type checker. Atlang defers type checking and translation
to these methods whenever an expression of type \textsf{string_in} is encountered.
